---
title: "Supervised Methods in Machine Learning report guideline: Exploration and Application of R for Data Science"
author: "Eduard Romero"
date: "2023-02-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic coding in R.

Please create a script in R to print the prime numbers from 1 to 100.

* Solution :
```{r results='hide'}

for (i in 1:100) {
  # Check if i is a prime number
  is_prime <- TRUE
  for (j in 2:(i-1)) {
    if (i %% j == 0) {
      is_prime <- FALSE
      break
    }
  }
  # Print the prime number
  if (is_prime && i > 1) {
    print(i)
  }
}
```
# Package installation and basic use of tidyverse.

From the website: https://r4ds.had.co.nz/transform.html,  solve and discuss in detail the following exercises:

* 5.2.4 Exercises: items 1, and 2
* 5.3.1 Exercises: all items
* 5.4.1 Exercises: items 2, 3, and 4
* 5.5.2 Exercises: items 1, and 2
* 5.6.7 Exercises: item 1
* 5.7.1 Exercises: item 2

* Solution 5.2.4 item 1: Find all flights that Had an arrival delay of two or more hours.
```{r results='hide'}
#Import data frame
library(nycflights13)
#Import library for commands
library(tidyverse)
#Import library for commands
library(dplyr)
#Call a tibble
flights
# Delay is a local varibale for use the command filter
Delay <- filter(flights, arr_delay >= 120 )
Delay

```
* Solution 5.2.4 item 2:    Find all flights that flew to Houston (IAH or HOU)

```{r results='hide'}
# fly_houston is local variable for use the command filter
fly_houston <- filter(flights, dest == "IAH" | dest == "HOU")
fly_houston

```
* Solution 5.3.1 item 1:How could you use arrange() to sort all missing values to the start?
```{r results='hide'}
# The operator pipeline %>$% is useful for concatenating multiple dplyr operations
flights %>% 
  # the arrange() function put na values last
  #  command desc sorted in descending order
    arrange(desc(is.na(dep_time)),
           desc(is.na(dep_delay)),
           desc(is.na(arr_time)), 
           desc(is.na(arr_delay)),
           desc(is.na(tailnum)),
           desc(is.na(air_time)))

```
* Solution 5.3.1 item 2: Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r results='hide'}
# when we use the arrange and desc functions we are making our table sort in descending order
arrange(flights, desc(dep_delay))
# when we use the arrange function only we are making our table sort in ascending order
arrange(flights,(dep_delay))

```
* Solution 5.3.1 item 3: Sort flights to find the fastest (highest speed) flights.
```{r results='hide'}
# When we use the arrange function only we are making our table sort in ascending order 
arrange(flights, (air_time))

```
* Solution 5.3.1 Item 4:Which flights traveled the farthest? Which traveled the shortest?
```{r results='hide'}
# when we use the arrange and desc functions we are making our table sort in descending order
arrange (flights, desc(distance))
# When we use the arrange function only we are making our table sort in ascending order 
arrange(flights, distance)

```
>[!NOTE]
   >
   > The function arrage is for order the vector ascending, Function arrage, desc is for order the vector descending

* Solution 5.4.1 item 2:What happens if you include the name of a variable multiple times in a select() call?

>[!NOTE]
   >
   > The function select() call ignores the duplication, Any duplicated variables are only included once
   
```{r results='hide'}
# when we use the select() function it will only show the variables mentioned within the table
select(flights, year, month,day,day,day)
```
* Solution 5.4.1 item 3:What does the any_of() function do? Why might it be helpful in conjunction with this vector? 

>[!NOTE]
   >
   > the function any_of() select variables contained in a character vector. They are especially useful for programming with selecting functions

```{r results='hide'}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, any_of(vars))
```
* Solution 5.4.1 item 4:Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

>[!NOTE]
   >
   > when we use this expression it will only show the variables what contains values of time, he default helper functions are insensitive to case. This can be changes by setting

```{r results='hide'}
select(flights, contains("TIME", ignore.case = FALSE))
```

* Solution 5.5.2 item 1:Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r results='hide'}
# Method with division integer and modulus
mutate(flights,
       dep_time = (dep_time %/% 100) * 60 + (dep_time %% 100),
       sched_dep_time = (sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100))
```
* Solution 5.2.2 item 2:Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r results='hide'}
flight_exercise <- select(flights, dep_time, arr_time, air_time)
mutate(flight_exercise, 
       arr_min = (arr_time %/% 100)*60 + (arr_time %% 100),
       dep_min = (dep_time %/% 100)*60 + (dep_time %% 100),
       fly_short = arr_min - dep_min)

```



* Solution 5.6.7 item 1: Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios

* A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
* A flight is always 10 minutes late.
* A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
* 99% of the time a flight is on time. 1% of the time it’s 2 hours late.


* Solution 5.7.1 item 1:Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.

*It works with groups. For example, you can use mutation to create new variables using only group statistics, or you can filter all groups or filter as usual.*

# Reporting with Rmarkdown
Discuss in detail 2 of the functions for data transformation explained in https://r4ds.had.co.nz/transform.html.
Discussion should be original (i.e. is not textually taken from any external source). It is expected that you at least:

* Use one chunk per explanation
* Change some of the chunk options and explain how the options affect the report.
    
    * Solution 1:
    
>[!NOTE]
   >
   > In this case we have created a dataframe with 4x4, using the mutate function we add another column where the professional performance of each minipak company worker will be evaluated.

```{r results='hold'}
#Load the library
library(dplyr)
# Creating a dataframe manual

Minipak <- data.frame(
  Employees_Name = c("Eduard Romero", "John Rodriguez", "Sergio Mendoza", "Andres Lote"),
    English_Level =c(6, 7, 5, 4),
    Jobs_Trips = c(1, 5, 2 ,1),
    Lvl_Engeenering = c(8, 10, 10, 10)
)
#Mutate Function
mutate(Minipak,AverageProfessional = English_Level+ Lvl_Engeenering*Jobs_Trips)
```
 * Solution 2:

```{r results='hold'}
#Load the library
library(dplyr)
# Creating a dataframe manual
UECCi <- data.frame(
  NameStudent =c("Juan","Eduard","Nicol","Daniela","Sergio", "David","Duvar", "Javier"),
  CareerName =c("Ing Mecatronica","Ing Electronica","Ing Mecanica","Ing Industrial","Ing Mecatronica", "Ing Electronica", "Lic Fisica","Ing Mecanico"),
   Score = c("3.5","3.0","4.0","4.5","3.8","3.1","3.9","4.0")
)
arrange(UECCi,Score  )
```
>[!NOTE]
   >
   > In this case we have created a dataframe, The arrange function allows us to organize our data by receiving the desired variable as a parameter and in ascending order.

# Chunks Options:

In this document we use two options of configurations, the first is {r results='Hide'}

The Second is {r results='hold'}
